General constraints on magic distillation protocols from statistical mechanics

### 2021-06-06 magic_paper.tex
added figures to show how Ha and Da look like for a>1 and a<1. Stuck at showing Ha > Hb wheneve a>b, otherwise I think we have all the results that we need. Need reaarangement and notation fixing.

We first define the classical $\alpha$-Renyi entropy in terms of the Wigner quasi-distribution of any state $\rho$ as
\begin{equation}
	H_\alpha(\rho) := \frac{1}{1-\alpha} \log \sum_{\bmx} W_{\rho}(\bmx)^\alpha,
\end{equation}
where $\alpha$ can take values of the form $2a / (2b+1)$, for non-negative integers $a,b$ to ensure that the argument of the logarithm is positive.
Note that for any quantum state $\rho$, $H_1(\rho)$ in particular is not well-defined  \nick{Notation clashes with quantum Renyi entropies}, while for any valid $\alpha$ we have that
\begin{equation}
	H_0(\rho) = 2\log_2 d = H_\alpha \left( \id/d \right),
\end{equation}
where $d$ is the dimension of $\rho$.
For $\rho = \rho_{\rm{S}}(\epsilon)$, we get
\begin{equation}
	H_{\alpha}\left(\rho_{\rm{S}}(\epsilon)\right) = \frac{1}{1-\alpha}\log_2 \left[ 8\left( \frac{1}{6} - \frac{1}{18}\epsilon \right)^\alpha + \left(-\frac{1}{3} + \frac{4}{9}\epsilon \right)^\alpha \right]
\end{equation}

\begin{figure}[h]
    \centering
    \includegraphics[scale=0.35]{figs/test/H_vs_eps_a.pdf}
    \caption{\ddd{[TO DO: Set $\epsilon = 0$ and determine region of $\alpha$ such that $H <0$.]} Here, the exponent is $\alpha =$ even$/$odd for $1.10 \leq \alpha \leq 1.50$. $H_\alpha$ is increasing in $\alpha$ and $\epsilon$ \nick{should be easy to show analytically}. First dashed contour from the right corresponds to $H_\alpha(\rho_{\rm{S}}(\epsilon)) = 0$. Specifically, $H_\alpha(\rho_{\rm{S}}(0)) = 0$ has a unique solution, which is the \nick{probably irrational} solution of $3^\alpha - 2^{3-\alpha} = 1$.
    }
    \label{fig:lcs}
\end{figure}

@@

Therefore, whenever $w_i \geq w_j$, we have that 
\begin{equation}
	-w_i^{\alpha-1} \leq -w_j^{\alpha-1},
\end{equation}
which leads to
\begin{equation}
	\frac{\partial H_\alpha}{\partial w_i} \leq \frac{\partial H_\alpha}{\partial w_j}.
\end{equation}
Therefore, $H_\alpha$ is Schur-concave and
\begin{equation}
	H_\alpha(\w) \leq H_\alpha(\w'),
\end{equation}
for any quasi-distributions $\w, \w'$ that obey $\w \succ \w'$.
\end{proof}
The set $F := \{2a/(2b-1): a,b \in \mathbb{N} \text{ and } a \geq b\}$ is dense in the rationals $\mathbb{Q}_{>1}$, which in turn are dense in the reals $\mathbb{R}_{>1}$, as any rational $c/d$ with $c>d$ can be approximated by a sequence $f_n := c2^n / (d2^n+1)$ with every element $f_n$ in the set $F$. \nick{can turn into lemma}

@@

\begin{theorem}\label{thm:Da_props} 
	Let $\tau$ be a state in the interior of $\F$ and $\alpha = 2a / (2b+1)$ for non-negative integers $a,b$
Then, the $\alpha$-Renyi divergence $D_\alpha(\rho, \tau)$ is well-defined for all states $\rho$, and the following hold:
\begin{enumerate}
\item $D_\alpha(\rho || \tau) \ge 0$ for all quantum states $\rho$.
\item  $D_\alpha(\rho || \tau) = 0$ if and only if $\rho =\tau$.
\item $D_\alpha(\rho^{\otimes n} || \tau^{\otimes n}) = n D_\alpha(\rho, \tau)$ for any $n \in \mathbb{N}$.
\item $D_\alpha(\rho || \tau) \geq D_\alpha (\E(\rho) || \E(\tau))$ for any free operation $\E$ such that $\E(\tau)$ is in the interior of $\F$.
\end{enumerate}
\end{theorem}
\begin{proof}
	Since $\tau$ is in the interior of $\F$, its Wigner function obeys $W_\tau(\x) >0$ for all points $\x$ in the phase space. 
In general, $W_\rho(\x)$ is a quasi-distribution, but the given form of $\alpha$ ensures that $W_{\rho}(\bmx)^\alpha \geq 0$. 
Therefore $D_\alpha (\rho || \tau)$ is always well-defined.

@@

\ddd{The following is a nice thing to establish.}
\begin{theorem} A state $\rho$ has Wigner negativity if and only if $H_\alpha (W_\rho (\x)) < 0$ for some $\alpha >1$.
\end{theorem}
Let's try to prove this.
\begin{proof} If $H_\alpha (W_\rho(\x) <0 $ then this implies that $\log \sum_{\x} W_\rho (\x)^\alpha > 1$, however it is known that $H_\alpha$ is always non-negative on probability distributions and therefore $W_\rho(\x)$ must be a strict quasi-distribution with negativity.

Conversely, suppose $\rho$ has negativity in its Wigner representation. This implies that $\sum_{\x} |W_\rho (\x)| >1$. We now consider $\sum_{\x} W_\rho (\x)^{a/b}$ for $a/b >1$ and $a$ even and $b$ odd. We have that
\begin{align}
\sum_{\x} W_\rho (\x)^{a/b} &= \sum_{\x} |W_\rho (\x)|^{a/b} \nonumber \\
&= \sum_{\x} |W_\rho (\x)|^{1+ \epsilon},
\end{align}
where $\epsilon = a/b - 1 >0$. By choosing $a$ and $b$ appropriately we can make $\epsilon$ arbitrarily close to $0$ and therefore there exists a sequence $\epsilon_n$ such that
\begin{equation}
\sum_{\x} |W_\rho (\x)|^{1+ \epsilon_n} \rightarrow \sum_{\x} |W_\rho (\x)| >1.
\end{equation}
Therefore, $H_\alpha (W_\rho (\x)) <0 $ at some point in the sequence, which completes the proof.
\end{proof}
Therefore Wigner negativity coincides with negativity in at least one Renyi entropy.
\ddd{[Looking at this limit it almost feels like the $\alpha \rightarrow 0^+$ limit is actually giving mana....modulo the technical problem of taking the limit without calculus. I will think more about this point.]}

>>>>

We first define the classical $\alpha$-Renyi entropy in terms of the Wigner quasi-distribution of any state $\rho$ as
\begin{equation}
	H_\alpha(\rho) := \frac{1}{1-\alpha} \log_2 \sum_{\bmx} W_{\rho}(\bmx)^\alpha,
\end{equation}
where $\alpha$ can take values of the form $2a / (2b+1)$, for non-negative integers $a,b$ to ensure that the argument of the logarithm is positive.
Note that for any quantum state $\rho$, $H_1(\rho)$ in particular is not well-defined  \nick{Notation clashes with quantum Renyi entropies}, while for any valid $\alpha$ we have that
\begin{equation}
	H_0(\rho) = 2\log_2 d = H_\alpha \left( \id/d \right),
\end{equation}
where $d$ is the dimension of $\rho$.
Inducing any magic state $\rho$ with depolarising noise, $\rho(\epsilon) = (1-\epsilon)\rho + \epsilon \sigma$ results in $H_{\alpha}\left(\rho(\epsilon)\right) \leq H_\alpha(\rho)$ because there exists a free operation $\E \in \O$ that transforms $\rho(\epsilon) = \E(\rho)$ and $H_\alpha$ is Schur-concave.

For $\rho = \rho_{\rm{S}}(\epsilon)$, we get
\begin{equation}
	H_{\alpha}\left(\rho_{\rm{S}}(\epsilon)\right) = \frac{1}{1-\alpha}\log_2 \left[ 8\left( \frac{1}{6} - \frac{1}{18}\epsilon \right)^\alpha + \left(-\frac{1}{3} + \frac{4}{9}\epsilon \right)^\alpha \right]
\end{equation}

\begin{figure}[h]
    \centering
    \includegraphics[scale=0.35]{figs/test/H_vs_eps_a.pdf}
    \caption{\ddd{[TO DO: Set $\epsilon = 0$ and determine region of $\alpha$ such that $H <0$.]} Here, the exponent is $\alpha =$ even$/$odd for $1.10 \leq \alpha \leq 1.50$. $H_\alpha$ is increasing in $\alpha$ and $\epsilon$. First dashed contour from the right corresponds to $H_\alpha(\rho_{\rm{S}}(\epsilon)) = 0$. Specifically, $H_\alpha(\rho_{\rm{S}}(0)) = 0$ has a unique solution, which is the \nick{probably irrational} solution of $3^\alpha - 2^{3-\alpha} = 1$.
    }
    \label{fig:H}
\end{figure}

\begin{figure}[h]
    \centering
    \includegraphics[scale=0.35]{figs/test/D_vs_eps_a.pdf}
    \caption{Here, the exponent is $\alpha =$ even$/$odd for $1.10 \leq \alpha \leq 1.50$. $D_\alpha$ is decreasing in $\alpha$ and $\epsilon$ (clear due to~\cref{thm:H2D}).
    }
    \label{fig:D}
\end{figure}

\begin{figure}[h]
    \centering
    \includegraphics[scale=0.35]{figs/test/H_test.pdf}
    \caption{Here, the exponent is $\alpha =$ even$/$odd for $0.10 \leq \alpha \leq 0.90$. $H_\alpha$ is decreasing in $\epsilon$ (not Schur-conocave for $\alpha < 1$) and increasing in $\alpha$ (clear due to~\cref{thm:H2D}).
    }
    \label{fig:Htest}
\end{figure}

\begin{figure}[h]
    \centering
    \includegraphics[scale=0.35]{figs/test/D_test.pdf}
    \caption{Here, the exponent is $\alpha =$ even$/$odd for $0.10 \leq \alpha \leq 0.90$. $D_\alpha$ is increasing in $\epsilon$ (not Schur-convex for $\alpha < 1$) and decreasing in $\alpha$ (clear due to~\cref{thm:H2D}).
    }
    \label{fig:Dtest}
\end{figure}

\begin{lemma}
	Let $\alpha \geq \beta > 1$ where $\alpha, \beta$ are even$/$odd.
	Then,
	\begin{equation}
		H_\alpha(\x) \geq H_\beta(\x) \text{ and } D_\alpha(\x || \r) \leq D_\beta(\x || \r)
	\end{equation}
	for any quasi-distribution $\x \in \mathbb{R}^N$ and probability distribution $\r \in \mathbb{R}^N$.
\end{lemma}
\begin{proof}
	For $\alpha \geq \beta > 1$, we have
	\begin{equation}
		0 > \frac{1}{1-\alpha} \geq \frac{1}{1-\beta},
	\end{equation}
	and $w_i^\alpha = |w_i|^\alpha \leq |w_i|^\beta = w_i^\beta$.

	\nick{how to proceed?}

	For any $\r$ with positive rational components, we have $D_\alpha(\x || \r) \leq D_\beta(\x || \r)$ due to~\cref{thm:H2D}.
\end{proof}

@@

Therefore, whenever $w_i \geq w_j$, we have that 
\begin{equation}
	-w_i^{\alpha-1} \leq -w_j^{\alpha-1},
\end{equation}
which leads to
\begin{equation}
	\frac{\partial H_\alpha}{\partial w_i} \leq \frac{\partial H_\alpha}{\partial w_j}.
\end{equation}
Therefore, $H_\alpha$ is Schur-concave and
\begin{equation}
	H_\alpha(\w) \leq H_\alpha(\w'),
\end{equation}
for any quasi-distributions $\w, \w'$ that obey $\w \succ \w'$.
\end{proof}
The set $F := \{2a/(2b-1): a,b \in \mathbb{N} \text{ and } a \geq b\}$ is dense in the rationals $\mathbb{Q}_{>1}$, which in turn are dense in the reals $\mathbb{R}_{>1}$, as any rational $c/d$ with $c>d$ can be approximated by a sequence $f_n := c2^n / (d2^n-1)$ with every element $f_n$ in the set $F$. \nick{can turn into lemma}

@@

\begin{theorem}\label{thm:Da_props} 
	Let $\tau$ be a state in the interior of $\F$ and $\alpha = 2a / (2b-1)$ for non-negative integers $a,b$ with $a \geq b$.
Then, the $\alpha$-Renyi divergence $D_\alpha(\rho, \tau)$ is well-defined for all states $\rho$, and the following hold:
\begin{enumerate}
\item $D_\alpha(\rho || \tau) \ge 0$ for all quantum states $\rho$.
\item  $D_\alpha(\rho || \tau) = 0$ if and only if $\rho =\tau$.
\item $D_\alpha(\rho^{\otimes n} || \tau^{\otimes n}) = n D_\alpha(\rho, \tau)$ for any $n \in \mathbb{N}$.
\item $D_\alpha(\rho || \tau) \geq D_\alpha (\E(\rho) || \E(\tau))$ for any free operation $\E$ such that $\E(\tau)$ is in the interior of $\F$.
\end{enumerate}
\end{theorem}
\begin{proof}
	Since $\tau$ is in the interior of $\F$, its Wigner function obeys $W_\tau(\x) >0$ for all points $\x$ in the phase space. 
In general, $W_\rho(\x)$ is a quasi-distribution, but the given form of $\alpha$ ensures that $W_{\rho}(\bmx)^\alpha \geq 0$. 
Therefore $D_\alpha (\rho || \tau)$ is always well-defined.

@@

\ddd{The following is a nice thing to establish.}
\begin{theorem} A state $\rho$ has Wigner negativity if and only if $H_\alpha (W_\rho (\x)) < 0$ for some $\alpha >1$.
\end{theorem}
Let's try to prove this.
\begin{proof} If $H_\alpha (W_\rho(\x)) <0 $ then this implies that $\log \sum_{\x} W_\rho (\x)^\alpha > 1$, and it is known that $H_\alpha$ is always non-negative on probability distributions, so $W_\rho(\x)$ must be a strict quasi-distribution with negativity.

Conversely, suppose $\rho$ has negativity in its Wigner representation. This implies that $\sum_{\x} |W_\rho (\x)| >1$. We now consider $\sum_{\x} W_\rho (\x)^{a/b}$ for $a/b >1$ and $a$ even and $b$ odd. We have that
\begin{align}
\sum_{\x} W_\rho (\x)^{a/b} &= \sum_{\x} |W_\rho (\x)|^{a/b} \nonumber \\
&= \sum_{\x} |W_\rho (\x)|^{1+ \epsilon},
\end{align}
where $\epsilon = a/b - 1 >0$. By choosing $a$ and $b$ appropriately we can make $\epsilon$ arbitrarily close to $0$ and therefore there exists a sequence $\epsilon_n$ such that
\begin{equation}
\sum_{\x} |W_\rho (\x)|^{1+ \epsilon_n} \rightarrow \sum_{\x} |W_\rho (\x)| >1.
\end{equation}
\nick{In this step, we also need to show that $H_\alpha$ is increasing in $\alpha$}
Therefore, $H_\alpha (W_\rho (\x)) <0 $ at some point in the sequence, which completes the proof.
\end{proof}
Therefore Wigner negativity coincides with negativity in at least one Renyi entropy.
\ddd{[Looking at this limit it almost feels like the $\alpha \rightarrow 1$ limit is actually giving mana....modulo the technical problem of taking the limit without calculus. I will think more about this point.]}
